Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-24 algorithm.tex
Added selfreferences and discussions

On the other side, parameter $\alpha$ determines when a region is considered of high
innovation. High values of $\alpha$ mean that a region will hardly be
elected as high-innovation, yielding less splitting operations. This
slows down zooming in on portions of configuration space and make the
distribution of the simulated configuration more uniform, as there are
fewer small regions with a high test configuration concentration. The
opposite happens when small values of $\alpha$ are used. In this
introductory work we are not explicitly focused on a deep
investigation on how these parameters can be fine tuned, nevertheless
we conducted some preliminary tests to confirm the relative stability
of $PS$ with different choices and to select some ``reasonable'' average
values to perform the design space exploration (see Figure 
\comment{TODO This is confirmed by results of \secR{parameter_impact}
}).

>>>>

On the other side, parameter $\alpha$ determines when a region is considered of high
innovation. High values of $\alpha$ mean that a region will hardly be
elected as high-innovation, yielding less splitting operations. This
slows down zooming in on portions of configuration space and make the
distribution of the simulated configuration more uniform, as there are
fewer small regions with a high test configuration concentration. The
opposite happens when small values of $\alpha$ are used. In this
introductory work we are not explicitly focused on investigating how
these parameters could be fine tuned, nevertheless
we conducted some preliminary tests to confirm the relative stability
of $PS$ with different choices and to select some ``reasonable'' average
values to perform the design space exploration (see for example Figure
TODO in Section \ref{sec:ee})
\comment{TODO This is confirmed by results of \secR{parameter_impact}
}).

### 2014-10-24 introduction.tex
AA: exploration vs exploitation; parallelism
Added selfreferences and discussions

@@

\cite{dellnitz2005covering} proposes three different methods for the numerical approximation of the Pareto set. The first two require the knowledge of the gradient of the objective functions, which is usually not available in embedded system design, since the closed form expression of the objective functions is not known a priori. The third one, called ``Sampling algorithm'', does not need this requirement and is the closest to ours, although the mechanism is different. At every iteration a set of regions of configuration space is constructed with the aim to cover the Pareto set. The size of these regions becomes smaller at each iteration so that the covering becomes tighter. A random set of test points are evaluated in each new region; only the regions containing non-dominated points will be considered in the next iteration. Our approach is different. First, in the Sampling algorithm only the selected regions are examined, even though there is no guarantee that all the Pareto points are included in them. As a consequence, the Sampling algorithm may erroneously and prematurely neglect some regions that may possibly contain a non-negligible number of Pareto points. Our algorithm avoids this, since even though a region has not provided non-dominated points, it is not neglected, rather it is simply populated with fewer test points, thus giving the chance to find non-dominated points in them. From this behavior also comes another important difference. The Sampling algorithm of \cite{dellnitz2005covering} keeps reinforcing the analysis of the regions selected in the previous iterations. In other words, it keeps zooming into the regions that have been considered ``interesting'', i.e. rich of non-dominated points, in the previous iterations. The philosophy behind our algorithm is different. A region, which has been considered interesting at a certain iterations, may be not in the future, since it has been thoroughly analyzed. On the other hand, a region, initially uninteresting, may turn out to be interesting in the future, once all the other regions have been already exploited. Therefore, from an iteration to the next, our algorithm zoom in on regions that are currently interesting and zoom out from the uninteresting ones,  spanning in a broader and more dynamical way the configuration space. In addition, in the Sampling region only the number of non-dominated points is accounted to see how interesting a region is. On the contrary, we also consider the \emph{innovation}, as already anticipated in the introduction.

Many studies, as \cite{coello_easmop} and others, focus on genetic approaches to solve multiobjective
optimization problems. Genetic approaches
have many advantages: they provide good accuracy, they are customizable and very general
and require no a-priori knowledge to the designer.  The strong
point of genetic approaches can be summarized as follows: they consist
in an exploration that is sufficiently broad (the mutation avoids being rigidly restricted to limited parts of parameter space), and, at the same time, not too scattered as it is guided by the performances of
the already evaluated configurations. In this work we will compare 
the proposed approach against a well known Multi-Objective Genetic
Algorithm~\cite{knowles_techrep06}.
We claim that the approach presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.

>>>>

A preliminary estimation of hardware/software dependencies in order to
reduce the design space is addressed in~\cite{Catania2008}, where
authors propose to deal with the software parameters separately,
statistically testing their average effect on objectives before the
actual exploration is performed. Also in~\cite{merging} authors
analyze the impact of including compilation parameters into the design
space space exploration. The results of both works confirm the
important role that hardware/software dependencies can play in a
multi-objective exploration and thus justify the choice of explicitly
including the two class of parameters in the design space without
making any distinction between them from the algorithm's perspective.

@@

Authors of \cite{dellnitz2005covering,dellnitz2009computation} propose
three different methods for the numerical approximation of the Pareto
set. The first two require the knowledge of the gradient of the
objective functions, which is usually not available in embedded system
design, since the closed form expression of the objective functions is
not known a priori. The third one, called ``Sampling algorithm'', does
not need this requirement and is the closest to ours, although the
mechanism is different. At every iteration a set of regions of
configuration space is constructed with the aim to cover the Pareto
set. The size of these regions becomes smaller at each iteration so
that the covering becomes tighter. A random set of test points are
evaluated in each new region; only the regions containing
non-dominated points will be considered in the next iteration. Our
approach is different. First, in the Sampling algorithm only the
selected regions are examined, even though there is no guarantee that
all the Pareto points are included in them. As a consequence, the
Sampling algorithm may erroneously and prematurely neglect some
regions that may possibly contain a non-negligible number of Pareto
points. Our algorithm avoids this, since even though a region has not
provided non-dominated points, it is not neglected, rather it is
simply populated with fewer test points, thus giving the chance to
find non-dominated points in them. From this behavior also comes
another important difference. The Sampling algorithm of
\cite{dellnitz2005covering} keeps reinforcing the analysis of the
regions selected in the previous iterations. In other words, it keeps
zooming into the regions that have been considered ``interesting'',
i.e. rich of non-dominated points, in the previous iterations. The
philosophy behind our algorithm is different. A region, which has been
considered interesting at a certain iterations, may be not in the
future, since it has been thoroughly analyzed. On the other hand, a
region, initially uninteresting, may turn out to be interesting in the
future, once all the other regions have been already exploited.
Therefore, from an iteration to the next, our algorithm zoom in on
regions that are currently interesting and zoom out from the
uninteresting ones,  spanning in a broader and more dynamical way the
configuration space. In addition, in the Sampling region only the
number of non-dominated points is accounted to see how interesting a
region is. On the contrary, we also consider the \emph{innovation}, as
already anticipated in the introduction.

Many studies, as \cite{coello_easmop} and others, focus on genetic
approaches to solve multiobjective optimization problems. Genetic
approaches have many advantages: they provide good accuracy, they are
customizable and very general and require no a-priori knowledge to the
designer. The strong point of genetic approaches can be summarized as
follows: they consist in an exploration that is sufficiently broad
(the mutation avoids being rigidly restricted to limited parts of
parameter space), and, at the same time, not too scattered as it is
guided by the performances of the already evaluated configurations.
Different variants of these algorithms have been proposed in
literature, including mixed approaches as in~\cite{Ascia2011382},
where authors specifically focuse on increasing the performance of a
evolutionaly exploration strategy, proposing a fuzzy based technique
to avoing the actual simulation of a certain number of configurations.
In particular, in this work we compare the proposed approach against a
well known Multi-Objective Genetic Algorithm~\cite{knowles_techrep06}
which is at best of our knowledge one of the most appropriate for the
particular scenario we adopted as testbed.  We claim that the approach
presented in this paper has most of the benefits of the genetic
approaches, although the rationale is completely different.
